---
title: "seabirds_final"
output:
  html_document:
    df_print: paged
---

```{r}
library(tidyverse)
library(janitor)
library(readxl)
library(here)
```

### Reading in the data
```{r}
clean_data <- read_csv("seabirds_cleaned_data.csv")
clean_data
```
# Breif

Clean the data for from the Sea bird observation to answer the questions 

# Assumptions 
Took out any recorded with "NO BIRDS RECORDED"

Took out sensu, lato (unidentified) or upper case letters at the end of common_name as these where not part of the name buy descriptions of the birds.

Have added Ext to the questions as not sure your definition of type and species ie Royal / Wandering albatross and Black-browed albatross are different birds or as there all the same species ie just count all as just albatross. My first answer is with the former, the Ext parts have them as the latter.

# Cleanning steps
Use a script file to clean

`full_join()` the 2 data sheets so not to loses any data

`clean_names()`

Renamed the columns for easier reading

`recode()` the data for easier reading

Removed descriptions of the birds form the common_name

Write to a .csv file


### Selecting the relevant data
```{r}
seabirds <- clean_data %>% 
              select(record_id, common_name, scientific_name, 
                     species_abbreviation, total_sighting, num_group_sighting, 
                     lat)
seabirds
```


### Questions

#### Q1 Which bird had the most individual sightings?
```{r}
seabirds %>% 
  group_by(common_name) %>% 
  summarise(count = sum(n())) %>% 
  arrange(desc(count))

# Wandering albatross	11293
```



#### Ext Q1 combining all the albatross as one type
```{r}
seabirds %>% 
  group_by(common_name) %>%
  filter(str_detect(common_name, "(?i)albatross")) %>% 
  ungroup() %>% 
  summarise(count = sum(total_sighting, na.rm = TRUE))

# All albatross 30424	
```



#### Q2 Which bird had the highest total count?
```{r}
seabirds %>% 
  group_by(common_name) %>%
  summarise(count = sum(total_sighting, na.rm = TRUE)) %>% 
  arrange(desc(count))

# Short-tailed shearwater	982553	
```



#### Ext Q2 combining all the shearwater as one type
```{r}
seabirds %>% 
  group_by(common_name) %>%
  filter(str_detect(common_name, "(?i)shearwater")) %>% 
  ungroup() %>% 
  summarise(count = sum(total_sighting, na.rm = TRUE))

# All shearwater 1394468		
```



#### Q3 Which bird had the highest total count above a latitude of -30?
```{r}
seabirds %>% 
  group_by(common_name) %>%
  filter(lat > -30) %>%
  summarise(count = sum(total_sighting, na.rm = TRUE)) %>%
  arrange(desc(count))
  

# Taking above a latitude of -30 as nearer the equator ie 0
# Wedge-tailed shearwater	855	
```



#### Ext Q3 combining all the shearwater as one type
```{r}
seabirds %>% 
  group_by(common_name) %>%
  filter(str_detect(common_name, "(?i)shearwater")) %>% 
  filter(lat > -30) %>%
  ungroup() %>% 
  summarise(count = sum(total_sighting, na.rm = TRUE))

# All shearwater seen above a latitude of -30 (nearer the the equator ie 0) 888		
```



#### Q4 How many different types of birds were only ever seen in groups of 1?
```{r}
seabirds %>% 
  group_by(common_name) %>% 
  mutate(max_sighting = max(num_group_sighting, na.rm = TRUE)) %>% 
  filter(max_sighting == 1) %>%
  distinct(common_name) 

# 23 birds only seen once
```


#### Ext Q4 If by types of birds you mean more individual species with no variations
```{r}
seabirds %>% 
  group_by(common_name) %>% 
  mutate(common_name = if_else(str_detect(common_name, 
                                          "(?i)shearwater"),"shearwater", 
                                                            common_name),
         common_name = if_else(str_detect(common_name, 
                                          "(?i)albatross"), "albatross",
                                                            common_name),
         common_name = if_else(str_detect(common_name, 
                                          "(?i)mollymawk"), "mollymawk",
                                                            common_name),
         common_name = if_else(str_detect(common_name, 
                                          "(?i)petrel"), "petrel",
                                                            common_name),
         common_name = if_else(str_detect(common_name, 
                                          "(?i)prion"), "prion",
                                                            common_name),
         common_name = if_else(str_detect(common_name, 
                                          "(?i)skua"), "skua",
                                                            common_name)
         ) %>% 
    mutate(max_sighting = max(num_group_sighting, na.rm = TRUE)) %>% 
  filter(max_sighting == 1) %>%
  distinct(common_name) 


# 2 individual species of birds only seen once, 
# when the variations are combined as one group
```




#### Q5 How many penguins were seen? (Hint: there are many types of penguin)
```{r}
seabirds %>% 
  group_by(common_name) %>%
  filter(str_detect(common_name, "(?i)penguin")) %>% 
  ungroup() %>% 
  summarise(count = sum(total_sighting, na.rm = TRUE))

# 158	penguins 
```




















